---
title: "Fundamentals of R, Block 3 Practical"
author: "Henrique Sposito and Livio Muller-Silva"
output:
  pdf_document: default
  html_document: default
subtitle: "Block 3 - Practical Visualizations"
---

# R Markdown: another way to store code

Markdown is a simple formatting syntax for authoring HTML, PDF, and Word documents.

Creating an R Markdown document is just like an R script, you just have to click the new document button and select R Markdowm from the options.

Markdown allows you to mix chunks of code (in light grey) with text and export a document with your code, text, and plots.

You can embed an R code chunk like this:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
 # knit options
options(scipen=999) # removes scientific notation
```

In the case above, we are just adjusting the setup for the document and loading some packages for our R Markdown document.

This is the best resource for information on [R Markdown](https://bookdown.org/yihui/rmarkdown/)!

*Tip: remember to always load the data at the beginning so your Rmarkdown will knit properly! Adding your data and Rmd files on the same folder also helps with path issues when knitting.*

Basics:

Section headers work with #:

# First-level header

## Second-level header

### Third-level header

For changing text styles use *:

*Italics*

**Bold**

***Italics and bold***

For inserting R code click on the **C** button above or use Cmd + Option + I on MAC (for Windows: Ctrl + Alt + I).

```{r character, echo=TRUE}
as.character("R Markdown is awesome")
```

Code chunks can be evaluated (should code be run?), included (should the code displayed in knitted document?), and much more. [rmarkdown](https://rmarkdown.rstudio.com/), as a [tidyverse](https://www.tidyverse.org/) package, also has a [cheat sheet](https://www.rstudio.com/resources/cheatsheets/)!

When you click the **Knit** button a document in HTML or PDF can be generated that includes both content as well as any embedded R code chunks within the document.

Lastly, R Markdown can be further used to create presentations in R (as the ones we use in class, see the [xaringan](https://bookdown.org/yihui/rmarkdown/xaringan.html) package) or even to write your Master's thesis (check out [iheidown](https://jhollway.github.io/iheiddown/)).

**If we want to knit this to PDF, we need to download a Latex engine first:**

```{r, eval=FALSE}
install.packages("tinytex")
tinytex::install_tinytex()
```

# Visualizations

## Setting up the Gapminder data

```{r}
library(gapminder) # load packages
gapminder <- gapminder::gapminder # create an object
summary(gapminder) # summary data
```

Before we start, the [ggplot2 book](https://ggplot2-book.org/introduction.html) is a great source for you to learn the details of visualizations in R (and the book was written using an R Markdown).

## Line plots: The evolution of life expectancy 

To create line plots in ggplot2 we use the [`geom_line()`](https://ggplot2.tidyverse.org/reference/geom_line.html) function.

[What are line plots good for?](https://rkabacoff.github.io/datavis/Time.html#TimeSeries)

Let's plot life expectancy in time, for Switzerland!

```{r}
library(ggplot2)
library(dplyr)
gapminder %>%
  filter(country == "Switzerland") %>% # filter for Switzerland
  ggplot(aes(x = year, y = lifeExp)) + # adds a first ggplot2 layer
  geom_line() # add a line
```

How can we improve this?

```{r}
gapminder %>%
  filter(country == "Switzerland") %>% # filter for Switzerland
  ggplot(aes(x = year, y = lifeExp)) + # adds a first ggplot2 layer
  geom_line() + # add a line
  labs(title = "Life expectancy in Switzerland from 1957 to 2007", # add title
       x = "Year",  # add label for x axis
       y = "Life expectation", # add label for y axis
       caption = "Source: Gapminder") # add caption
```

Do you think that life expectancy increased for across all continents in time?

```{r message=FALSE, warning=FALSE}
gapminder %>%
  group_by(continent, year) %>% # group by year and country
  summarise(Avg_life_expectancy = mean(lifeExp)) %>%  #  mean life expectancy
  ggplot(aes(x = year, y = Avg_life_expectancy, aes(color = continent))) + # map
  geom_line(aes(color = continent)) +
  geom_point(aes(color = continent)) + #here we are mapping color by continent at the geom_level
  labs(title = "Life expectancy by continent from 1957 to 2007", # add title
       x = "Year", # add
       y = "Average life expectation", # add lable for y axis
       caption = "Source: Gapminder") + # add caption
  theme_minimal() # add theme
# What if I wanted to add a line for the average here, as above?
```

Could you make the same line plots for GDP per capita across continents in time?

## Bar plots: Life expectancy from 1957 to 2007 across continents 

To create bar plot in ggplot2 we use the [`geom_col()`](https://ggplot2.tidyverse.org/reference/geom_col.html) function.

[What are bar plots good for?](https://www.data-to-viz.com/graph/barplot.html)

Let's start with a simple bar plot, one continent at two different points in time.

```{r}
gapminder %>% # select data
  filter(year == 1957 | year == 2007,
         continent == "Oceania") %>% # filter for years of interest and Oceania
  ggplot(aes(x = country, # country in the x axis
             y = lifeExp, # map average life expectancy in the y axis
             fill = as.factor(year))) +
  # map fill by year (as factor), fill means the color filling the geom...
  geom_col()
```

What is the issue with this plot?

```{r}
gapminder %>% # select data
  filter(year == 1957 | year == 2007,
         continent == "Oceania") %>% # filter for years of interest and Oceania
  ggplot(aes(x = country, # country in the x axis
             y = lifeExp, # map average life expectancy in the y axis
             fill = as.factor(year))) +
  # map fill by year (as factor), remember that fill means the color filling the geom
  geom_col(position= "dodge") # separated the columns
```

This looks nicer, but there are several improvements we can still make!

```{r}
gapminder %>% # select data
  filter(year == 1957 | year == 2007,
         continent == "Oceania") %>% # filter for years of interest and Oceania
  ggplot(aes(x = country, # country in the x axis
             y = lifeExp, # map average life expectancy in the y axis
             fill = as.factor(year))) +
  # map fill by year (as factor), remember that fill means the color filling the geom
  geom_col(position= "dodge") + # separated the columns
  labs(title = "Life expectancy in 1957 and in 2007 in Oceania", # add a title
       x = "Country", # add a label for x axis
       y = "Life Expectancy", # add a label for y axis
       fill = "Year") + # sub legend for fill
  scale_fill_manual(values = c("darkgreen", "grey")) + # manually set colors
  theme(title = element_text(color= "gold")) 
  # adds a theme; theme_classic is a "clean" theme that removes unnecessary stuff
```

Lastly, let's use the same bar plots the average difference in life expectancy from 1957 to 2007 across continents.

```{r, message=FALSE, warning=FALSE}
gapminder %>% # select data
  filter(year == 1957 | year == 2007) %>% # filter for years of interest
  group_by(continent, year) %>% # group by year and country
  summarise(Avg_life_expectancy = mean(lifeExp)) %>% # get the means by the groups
  ggplot(aes(x = continent, # map continent in the x axis
             y = Avg_life_expectancy, # map average life expectancy in the y axis
             fill = as.factor(year))) + # fill mapping by year
  geom_col(position = "dodge") + # defining the position of stat in bars
  labs(title = "Life expectancy from 1957 to 2007", # add a title
       x = "Continent", # add a label for x axis
       y = "Average Life Expectancy", # add a label for y axis
       fill = "Year") + # sub legend for fill
  scale_fill_manual(values = c("gray", "darkgreen")) + # manually set colors
  theme_classic() # add a theme
```

## Box plots: Distribution of life expectancies across continent

To create scatter plots in ggplot2 we use the [`geom_boxplot()`](https://ggplot2.tidyverse.org/reference/geom_boxplot.html) function.

[What are box plots good for?](https://www.r-bloggers.com/2021/11/how-to-make-stunning-boxplots-in-r-a-complete-guide-with-ggplot2/)

```{r}
gapminder %>% # get data
  filter(year == 1957) %>% # filter by year
  ggplot(aes(x = continent, y = lifeExp, fill = continent)) + # map first layer
  geom_boxplot() + # box plot
  labs(title = "Life expectation by continent in 1957", # adds title
       x = "Continent", # adds x axis label
       y = "Life expectation", # adds y axis label
       caption = "Source: Gapminder") # adds caption
```

There is a lot of redundant information in this plot, no?

```{r}
library(RColorBrewer)
display.brewer.all(colorblindFriendly = TRUE, type = "qual")
# "qual" stands for qualitative!
gapminder %>% 
  filter(year == 1957 | year == 2007) %>% # filter
  ggplot(aes(x = continent, y = lifeExp, fill = continent)) + # fill by continent
  geom_boxplot() + # box plot
  facet_wrap(~year) + # separate plots by year; 2 different plots in this case
  labs(title = "Life expectation by continent in 1957 and 2007",
       x = "", # remove labels for x axis
       y = "Life expectation",
       caption = "Source: Gapminder") +
  theme(legend.position="none") +
  scale_fill_brewer(palette = "Set2") 
#scale_fill_brewer() colours the aesthetics fill with the "Set2"
```

Could you make the same box plot for GDP?

## Scatter plots: Population, life expectancy and GDP

To create scatter plots in ggplot2 we use the [`geom_point()`](https://ggplot2.tidyverse.org/reference/geom_point.html) function.

[What are scatter plots good for?](https://chartio.com/learn/charts/what-is-a-scatter-plot/)

```{r}
library(ggthemes) # This package has several aditional cool themes!
gapminder %>% 
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(size = pop, color = continent)) +
  # size points by population and color by continent
  geom_smooth(se = FALSE, color = "gray") + # add a smoothed line
  labs(title = "How much life can money buy?",
       x = "GDP per capita", 
       y = "Life expectation",
       caption = "Source: Gapminder") + # add caption
  scale_x_log10() +
  theme_stata() # add theme from the ggthemes package
```
