---
title: "Swearing and Religion in American Rap in the 21st Century"
author: "Henrique Sposito and Livio Silva-Muller"
date:
output:
  pdf_document: default
  html_document: default
---
# Data

## Loading data

As a case study, let's analyze and compare a few famous American Rappers in the 21st Century.

```{r setup, include=FALSE}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
# removes scientific notation
options(scipen=999)
# load packages
library(dplyr)
library(stringr)
library(tm)
library(ggplot2)
library(tidyr)
library(plotly)
```

```{r data, include=FALSE}
# Your path to data will be different here, please change accordingly!
setwd("~/Documents/GitHub/Fundamentals_of_R_IHEID2022/Fundamentals_workshop_2023/Case Study/American Rappers")
Kanye_West <- readRDS("Kanye_West.RDS")
Jay_Z <- readRDS("Jay_Z.RDS")
Eminem <- readRDS("Eminem.RDS")
Kendrick_Lamar <- readRDS("Kendrick_Lamar.RDS")
Nicki_Minaj<- readRDS("Nicki_minaj.RDS")
Cardi_B<- readRDS("Cardi_B.RDS")
Queen_latifah <- readRDS("Queen_latifah.RDS")
```

After loading the packages and datasets, we need to join these together.

If we simply join the datasets though, we will lose the name of the artist. So let's add a name variable to each dataset before joining them.

```{r names}
Eminem$rapper <- "Eminem"
Kanye_West$rapper <- "Kanye West"
Jay_Z$rapper <- "Jay Z"
Kendrick_Lamar$rapper <- "Kendrick Lamar"
Queen_latifah$rapper <- "Queen Latifah"
Cardi_B$rapper <- "Cardi B"
Nicki_Minaj$rapper <- "Nicki Minaj"
# # How could you do this in tidy?
# Eminem <- Eminem %>% mutate( rapper= "Eminem")
```

## Joining data

Now let's join data!

```{r join, message=FALSE, warning=FALSE}
american_rappers <- full_join(Eminem, 
                              full_join(Kanye_West,
                                        full_join(Jay_Z, 
                                                  full_join(Nicki_Minaj,
                                                                   full_join(Cardi_B, 
                                                                             full_join(Queen_latifah,Kendrick_Lamar))))))
```

We can only join them without specifying the "by =" argument because the datasets have the same variables with the same names.
However, this is not a very elegant solution. A more elegant solution for joining multiple datasets would entail using the ´{purrr}´ package.

```{r alternative join, eval=FALSE}
#library(purrr)
#american_rappers <- list(Eminem, Kanye_West, Jay_Z, Kendrick_Lamar,Nicki_Minaj, Cardi_B, Queen_latifah) %>% reduce(full_join)
```


## Data cleaning and wrangling

Let's clean and wrangle the data now.

First, there are some songs in the dataset that do not come from rappers' albums, but from somewhere else.
In the album variables in the dataset, songs that come from an album from one of the rappers start with "album:".

```{r album}
american_rappers$album <- ifelse(startsWith(american_rappers$album, "album:"),
                                 gsub("album:", "", american_rappers$album),
                                 NA_character_)
# # The same with tidy:
# american_rappers <- american_rappers %>%
#   mutate(album = ifelse(stringr::str_detect(album, "album:"),
#                         stringr::str_replace(album, "album:", ""),
#                         NA_character_))
```

Second, we can also extract the date from the album varible to create an 4 digit year variable.

```{r year}
american_rappers$year <- as.numeric(stringr::str_extract_all(american_rappers$album,
                                                             "[:digit:]{4}"))
```

Third, let's remove songs for which we are missing the album or that come from album collaborations.

```{r na rm}
american_rappers <- na.omit(american_rappers)
```

Fourth, let's clean the text by removing signs, transform to lower case, and more.

```{r clean text}
#removing punctuation
american_rappers$lyrics <- tm::removePunctuation(american_rappers$lyrics)
#making all lowercase
american_rappers$lyrics <- tolower(american_rappers$lyrics)
# remove line markers
american_rappers$lyrics <- gsub("\r|\n", " ", american_rappers$lyrics)
# remove punctuation
american_rappers$title <- tm::removePunctuation(american_rappers$title)
# remove the years from albums
american_rappers$album <- stringr::str_remove_all(american_rappers$album,
                                                  "\\([:digit:]{4}\\)")
```

# Analysis

## Dictionary

Create a dictionary for religion and swearing: can you help?

```{r dictionary}
swear_words <- "fuck|bitch|pussy|shit|dick|ass|cunt"
religious_words <- "god|bible|jesus|hell|heaven|lord|praise"
```

## Count

```{r count}
# base for counting swear words
american_rappers$swear_words <- stringr::str_count(american_rappers$lyrics,
                                                   swear_words)
# tidy for religion
american_rappers <- american_rappers %>%
  mutate(religious_words = stringr::str_count(american_rappers$lyrics, 
                                              religious_words))
```

## Swearing in American Rappers' songs

```{r swearing in songs, message=FALSE, warning=FALSE}
ggplot(american_rappers, aes(year, swear_words)) +
  geom_point(aes(size = swear_words), color="gold") +
  geom_text(aes(label = title), check_overlap=T, size=3) +
  labs(x = "", y = "Count",
       title = "Swearing in American Rappers' songs",
       subtitle= "989 songs by seven American Rappers since 1990")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey",
                                        size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none")
```

What are some problems with this plot? How can we deal with outliers?

```{r swear words outliers, message=FALSE, warning=FALSE}
american_rappers %>% filter(swear_words < 100) %>%
  ggplot(., aes(year, swear_words)) +
  geom_point(aes(color=rapper))+
  geom_text(aes(label = title), check_overlap=T, size=3, face="bold") +
  scale_color_brewer(palette="Set2")+
  labs(x = "", y = "Count",
       title = "Swearing in American Rappers' songs",
       subtitle= "989 songs by seven American Rappers since 1990")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

```{r swear words outliers nicer, message=FALSE, warning=FALSE}
american_rappers %>% filter(swear_words < 100) %>%
  ggplot(., aes(year, swear_words)) +
  geom_text(aes(label = title, color=rapper), check_overlap=T, size=3, face="bold") +
  scale_color_brewer(palette="Set2")+
  labs(x = "", y = "Count",
       title = "Swearing in American Rappers' songs",
       subtitle= "989 songs by seven American Rappers since 1990")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

```{r swearing in albums, include=FALSE, message=FALSE, warning=FALSE}
# Want to see swearing by albums?
american_rappers %>% group_by (album, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE))%>%
  ggplot(., aes(x=swear_words, y=reorder(album, swear_words))) +
  geom_bar(aes(fill=rapper), stat="identity") +
  scale_fill_brewer(palette="Set2")+
  labs(x = "Count", y = "Album Title",
       title = "Swearing in American Rappers' Albums",
       subtitle= "61 Albums since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

## Has swearing increased, or not, in time

```{r swear time, message=FALSE, warning=FALSE}
american_rappers %>% group_by (year, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE))%>%
  ggplot(., aes(year, swear_words)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "Year", y = "Swear Words Count",
       title = " Swear words by year",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

What are some issue is with this analysis?

Yes, we need to normalize scores!

But what would be the best normalization given our data?

```{r swearing in time normalized year, message=FALSE, warning=FALSE}
american_rappers %>%
  group_by(year, rapper) %>%
  mutate(songs_per_year = n()) %>%
  group_by(year,songs_per_year, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE)) %>%
  mutate(normalized_swear_words = swear_words/songs_per_year) %>% 
  #all lines until here are normalizing by songs per year
  ggplot(., aes(year, normalized_swear_words)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "", y = "Normalized count by songs per year",
       title = " Average swearing per song by year",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

Let's actually normalize by the number of words per song

```{r swearing per year normalized, message=FALSE, warning=FALSE}
american_rappers$word_per_song <-str_count(american_rappers$lyrics, "\\w+")
# counting words in lyrics
american_rappers$normalized_swear_words2 <- american_rappers$swear_words /
  american_rappers$word_per_song # normalize for swear words
american_rappers$normalized_religious_words2 <- american_rappers$religious_words / 
  american_rappers$word_per_song # normalize for religious words
# Plot
american_rappers %>%
  ggplot(., aes(year, normalized_swear_words2)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "", y = "Normalized Count (by song length)",
       title = " Average number of swear words per song",
       subtitle= "782 songs since 1996") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

What happens with the unit of measurement (i.e. average number of swear words per song)?

## Religious vocabulary in time: smoothing is okay?

```{r religion smoothed, message=FALSE, warning=FALSE}
american_rappers %>%
  group_by(year) %>%
  summarise(normalized_s=mean(normalized_swear_words2),
            normalized_r=mean(normalized_religious_words2)) %>%
    pivot_longer(2:3, names_to="topic", values_to="normalized_count2") %>%
  ggplot(., aes(x=year, y=normalized_count2, color=topic)) +
  geom_smooth(se=FALSE)+
      scale_color_brewer(palette="Set2")+
  labs(x = "", y = "Average number of swear words per song",
       title = " Religion and Swearing in American Rappers' Repertoire",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

```{r religion line, message=FALSE, warning=FALSE}
american_rappers %>%
  group_by(year) %>%
  summarise(normalized_s=mean(normalized_swear_words2),
            normalized_r=mean(normalized_religious_words2)) %>%
    pivot_longer(2:3, names_to="topic", values_to="normalized_count2") %>%
  ggplot(., aes(x=year, y=normalized_count2, color=topic)) +
  geom_line(size=1)+
    scale_color_brewer(palette="Set2")+
  labs(x = "", y = "Average number of swear words per song",
       title = " Religion and Swearing in American Rappers' Repertoire",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

Do you think there is a trend a wider trend in American Rappers using more religious words in songs?

```{r religion by rapper, message=FALSE, warning=FALSE}
american_rappers %>%
  pivot_longer(9:10, names_to= "topic", values_to="normalized_count2") %>%
  ggplot(., aes(x=year, y=normalized_count2, color=topic)) +
  geom_smooth(se=FALSE) +
    scale_color_brewer(palette="Set2")+
  labs(x = "", y = "Smoothed average number of swear words per song",
       title = " Religion and Swearing in American Rappers' Repertoire",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom") +
  facet_wrap(~rapper)
```

```{r regression variables and models}
american_rappers <- american_rappers %>%
  mutate(gender=case_when(rapper == "Cardi B" ~ "Female",
                          rapper =="Nicki Minaj" ~ "Female",
                          rapper == "Queen Latifah" ~"Female", TRUE ~ "Male"),
         two_thousands=ifelse(year>=2000, "Before 2000s","After 2000s"))

library(plm)

model <- plm(normalized_swear_words2 ~ normalized_religious_words2 + as.factor(gender) + word_per_song+
            as.factor(two_thousands),
   	                  data = american_rappers,
                     	model = "within", index = "rapper")
```

Is there really a wider trend or this is all driven by Kanye?

## The Kardashian effect

Is there a Kardashian effect in this switch we see for Kayne?

Kanye and Kim were a couple from 2011 to 2020, let's create a Kim variable!

```{r kim var}
Kanye_West <- american_rappers %>% 
  filter(rapper== "Kanye West")%>%
  mutate(kim_kardashian= case_when(year > 2010 & year <= 2020  ~ "With Kim",
                                   year < 2011  ~ "Before Kim",
                                   year >2019  ~ "After Kim"))
```

```{r kim effect reordered, message=FALSE, warning=FALSE}
Kanye_West$kim_kardashian <- factor(Kanye_West$kim_kardashian,                      
                                    levels = c("Before Kim",
                                               "With Kim",
                                               "After Kim"))
Kanye_West %>%
  gather("topic", "word_count", 6:7) %>%
  group_by(year) %>%
  mutate(songs_per_year = n()) %>%
  group_by(songs_per_year, topic, kim_kardashian) %>%
  summarise(word_count = sum(word_count, na.rm = TRUE))%>%
  mutate(normalized_word_count = word_count/songs_per_year)%>%
  ggplot(., aes(x=kim_kardashian, y=normalized_word_count, fill=topic)) +
  geom_bar(stat="identity", position="dodge") +
  labs(x = "", y = "Average swearing per song",
       title = " The Kardashian Effect?",
       subtitle= "214 songs in 13 albums since 2004.")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

## Lamar's Repertoire in words

Let's do a basic word frequency analysis.

```{r lamar cleaning, include=FALSE, message=FALSE, warning=FALSE}
# The function "function()" creates a function; in this case, our function cleans the text in multiple ways, by applying already existing functions from tm.

cleanCorpus<-function(corpus, customStopwords){
  corpus <- tm_map(corpus, content_transformer(qdapRegex::rm_url))
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, removeWords, customStopwords)
  return(corpus)
}

stops <- c(stopwords('SMART'), "dont", "youre")
# stopwords ('SMART') is a vector with common words (a, the, about...) that we want to remove from the songs

Kendrick_Lamar <- american_rappers %>%
  filter(rapper=="Kendrick Lamar") # creating a dataset only for Kendrick
  
# The next few lines perform a few operations to create a dataset counting appearances of words in all songs.
lamarCorpus_t <- VCorpus(VectorSource(Kendrick_Lamar$lyrics))
lamarCorpus_t <- cleanCorpus(lamarCorpus_t, stops)
LamarTDM_t  <- TermDocumentMatrix(lamarCorpus_t)
lamar_dtm_t  <- DocumentTermMatrix(lamarCorpus_t)

# create a document term matrix
LamarTDMm_t <- as.matrix(LamarTDM_t)
LamarSums_t <- rowSums(LamarTDMm_t)
LamarFreq_t <- data.frame(word=names(LamarSums_t),frequency=LamarSums_t)
rownames(LamarFreq_t) <- NULL
topWords_t  <- subset(LamarFreq_t, LamarFreq_t$frequency >= 50)
#here we are subsetting to those that appear more than 50 times.

topWords_t  <- topWords_t[order(topWords_t$frequency, decreasing=F),]
topWords_t$word <- factor(topWords_t$word, levels=unique(as.character(topWords_t$word)))
# getting the top words in frequency
```

```{r lamar plot, message=FALSE, warning=FALSE}
ggplot(topWords_t, aes(x=word, y=log(frequency))) +
  geom_bar(stat="identity", fill='gold') +
  coord_flip()+
  geom_text(aes(label=frequency), colour="black",hjust=1.25, size=3.0)+
  labs( x="", y= "", title= "Most frequent words in Lamar's repertoire",
        subtitle="All albums since 2011")+
  theme(panel.background = element_rect ("white", "black", .5, "solid"),
        panel.grid.major = element_line(color="grey", size=0.3, linetype= "solid"),
        axis.text = element_text(color="black", size=10),
        title = element_text(color="black", size=10, face="bold"),
        axis.text.x=element_blank(),
        plot.subtitle = element_text(color="black", size=9, face= "plain"),
        legend.position = "bottom")
```

Is our sample of swear/religious word bias?

Do you think this is the same for the other rappers in the sample?
