---
title: "Swearing and Religion in American Rap in the 21st Century"
author: "Henrique Sposito and Livio Silva-Muller"
date: '2022-10-14'
output:
  pdf_document: default
  html_document: default
---

# Data

## Loading data

As a case study, let's analyze and compare a few famous American Rappers in the 21st Century.

```{r setup, include=FALSE}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
# removes scientific notation
options(scipen=999)
# load packages
library(dplyr)
library(stringr)
library(tm)
library(ggplot2)
library(tidyr)
library(plotly)
```

```{r data, include=FALSE}
# Your path to data will be different here, please change accordingly!
Kanye_West <- readRDS("~/Documents/GitHub/Fundamentals_of_R_IHEID2022/Lecture 3/Class Material/Kanye_West.RDS")
Jay_Z <- readRDS("~/Documents/GitHub/Fundamentals_of_R_IHEID2022/Lecture 3/Class Material/Jay_Z.RDS")
Eminem <- readRDS("~/Documents/GitHub/Fundamentals_of_R_IHEID2022/Lecture 3/Class Material/Eminem.RDS")
Kendrick_Lamar <- readRDS("~/Documents/GitHub/Fundamentals_of_R_IHEID2022/Lecture 3/Class Material/Kendrick_Lamar.RDS")
```

After loading the packages e datasets, we need to join these together.

If we simply join the datasets though, we will lose the name of the artist. So let's add a name variable to each dataset before joining them.

```{r names}
Eminem$rapper <- "Eminem"
Kanye_West$rapper <- "Kanye West"
Jay_Z$rapper <- "Jay Z"
Kendrick_Lamar$rapper <- "Kendrick Lamar"
# # How could you do this in tidy?
# Eminem <- Eminem %>% mutate( rapper= "Eminem")
```

## Joining data

Now let's join data!

```{r join, message=FALSE, warning=FALSE}
american_rappers <- full_join(Eminem, 
                              full_join(Kanye_West,
                                        full_join(Jay_Z, Kendrick_Lamar)))
```

Please notice that we can only join them without specifying the "by =" argument because the datasets have the same variables with the same names.

However, this is not a very elegant solution.
A more elegant solution for joining multiple datasets would entail using the ´{purrr}´ package.

```{r alternative join, eval=FALSE}
library(purrr)
american_rappers <- list(Eminem, Kanye_West, Jay_Z, Kendrick_Lamar) %>%  reduce(full_join)
```

## Data cleaning and wrangling

Let's clean and wrangle the data now.

First, there are some songs in the dataset that do not come from rappers' albums, but from somewhere else.
In the album variables in the dataset, songs that come from an album from one of the rappers start with "album:".

```{r album}
american_rappers$album <- ifelse(startsWith(american_rappers$album, "album:"),
                                 gsub("album:", "", american_rappers$album),
                                 NA_character_)
# # The same with tidy:
# american_rappers <- american_rappers %>%
#   mutate(album = ifelse(stringr::str_detect(album, "album:"),
#                         stringr::str_replace(album, "album:", ""),
#                         NA_character_))
```

Second, we can also extract the date from the album varible to create an 4 digit year variable.

```{r year}
american_rappers$year <- as.numeric(stringr::str_extract_all(american_rappers$album,
                                                             "[:digit:]{4}"))
```

Third, let's remove songs for which we are missing the album or that come from album collaborations.

```{r na rm}
american_rappers <- na.omit(american_rappers)
```

Fourth, let's clean the text by removing signs, transform to lower case, and more.

```{r clean text}
#removing punctuation
american_rappers$lyrics <- tm::removePunctuation(american_rappers$lyrics)
#making all lowercase
american_rappers$lyrics <- tolower(american_rappers$lyrics)
# remove line markers
american_rappers$lyrics <- gsub("\r|\n", " ", american_rappers$lyrics)
# remove punctuation
american_rappers$title <- tm::removePunctuation(american_rappers$title)
# remove the years from albums
american_rappers$album <- stringr::str_remove_all(american_rappers$album,
                                                  "\\([:digit:]{4}\\)")
```

# Analysis

## Dictionary

Create a dictionary for religion and swearing: can you help?

```{r dictionary}
swear_words <- "fuck|bitch|pussy|shit|dick|ass|cunt"
religious_words <- "god|bible|jesus|hell|heaven|lord|praise"
```

## Count

```{r count}
# base for counting swear words
american_rappers$swear_words <- stringr::str_count(american_rappers$lyrics,
                                                   swear_words)
# tidy for religion
american_rappers <- american_rappers %>%
  mutate(religious_words = stringr::str_count(american_rappers$lyrics, 
                                              religious_words))
```

## Swearing in American Rappers' songs

```{r swearing in songs, message=FALSE, warning=FALSE}
ggplot(american_rappers, aes(year, swear_words)) +
  geom_point(aes(size = swear_words), color="gold") +
  geom_text(aes(label = title), check_overlap=T, size=3) +
  labs(x = "", y = "Count",
       title = "Swearing in American Rappers' songs",
       subtitle= "782 songs from Eminem, Jay Z, Kendrick Lamar, and Kanye West.")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey",
                                        size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none")
```

What are some problems with this plot? How can we deal with outliers?

```{r swear words outliers, message=FALSE, warning=FALSE}
american_rappers %>% filter(swear_words < 100) %>%
  ggplot(., aes(year, swear_words)) +
  geom_point(aes(color = rapper)) +
  geom_text(aes(label = title), check_overlap=T, size=3) +
  labs(x = "", y = "Count",
       title = "Swearing in American Rappers' songs",
       subtitle= "782 songs from Eminem, Jay Z, Kendrick Lamar, and Kanye West.")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

```{r swearing in albums, include=FALSE, message=FALSE, warning=FALSE}
# Want to see swearing by albums?
american_rappers %>% group_by (album, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE))%>%
  ggplot(., aes(x=swear_words, y=reorder(album, swear_words))) +
  geom_bar(aes(fill=rapper), stat="identity") +
  labs(x = "Count", y = "Album Title",
       title = "Swearing in American Rappers' Albums",
       subtitle= "46 Albums since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

## Has swearing increased, or not, in time

```{r swear time, message=FALSE, warning=FALSE}
american_rappers %>% group_by (year, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE))%>%
  ggplot(., aes(year, swear_words)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "Year", y = "Swear Words",
       title = " Swear words by year",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

What are some issue is with this analysis?

Yes, we need to normalize scores!

But what would be the best normalization given our data?

```{r swera in time normalized year, message=FALSE, warning=FALSE}
american_rappers %>%
  group_by(year, rapper) %>%
  mutate(songs_per_year = n()) %>%
  group_by(year,songs_per_year, rapper) %>%
  summarise(swear_words = sum(swear_words, na.rm = TRUE)) %>%
  mutate(normalized_swear_words = swear_words/songs_per_year) %>% 
  #all lines until here are normalizing by songs per year
  ggplot(., aes(year, normalized_swear_words)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "", y = "Normalized count by songs per year",
       title = " Average swearing per song by year",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

Let's actually normalize by the number of songs in per album.

```{r swera per year normalized, message=FALSE, warning=FALSE}
american_rappers$word_per_song <-str_count(american_rappers$lyrics, "\\w+")
# counting words in lyrics
american_rappers$normalized_swear_words2 <- american_rappers$swear_words /
  american_rappers$word_per_song # normalize for swera words
american_rappers$normalized_religious_words2 <- american_rappers$religious_words / 
  american_rappers$word_per_song # normalize for religious words
# Plot
american_rappers %>%
  ggplot(., aes(year, normalized_swear_words2)) +
  geom_smooth(se=FALSE, color="black") +
  labs(x = "", y = "Normalized Count (by song length)",
       title = " Average number of swear words per song",
       subtitle= "782 songs since 1996") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none") +
  facet_wrap(~rapper)
```

What happens with the unit of measurement (i.e. average number of swear words per song)?

## Religious vocabulary in time

```{r religion, message=FALSE, warning=FALSE}
american_rappers %>%
  gather("topic", "normalized_count2", 9:10) %>%
  ggplot(., aes(x=year, y=normalized_count2, color=topic)) +
  geom_smooth(se=FALSE) +
  labs(x = "", y = "Average number of swear words per song",
       title = " Religion and Swearing in American Rappers' Repertoire",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

Do you think there is a trend a wider trend in American Rappers using more religious words in songs?

```{r religion by rapper, message=FALSE, warning=FALSE}
american_rappers %>%
  gather("topic", "normalized_count2", 9:10) %>%
  ggplot(., aes(x=year, y=normalized_count2, color=topic)) +
  geom_smooth(se=FALSE) +
  labs(x = "", y = "Average number of swear words per song",
       title = " Religion and Swearing in American Rappers' Repertoire",
       subtitle= "782 songs since 1996")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom") +
  facet_wrap(~rapper)
```

Is there really a wider trend or this is all driven by Kayne?

## The Kardashian effect

Is there a Kardashian effect in this switch we see for Kayne?

Kanye and Kim were a couple from 2011 to 2020, let's create a Kim variable!

```{r kim var}
Kanye_West <- american_rappers %>% 
  filter(rapper== "Kanye West")%>%
  mutate(kim_kardashian= case_when(year > 2010 & year <= 2020  ~ "With Kim",
                                   year < 2011  ~ "Before Kim",
                                   year >2019  ~ "After Kim"))
```

```{r kim effect, message=FALSE, warning=FALSE}
Kanye_West %>%
  gather("topic", "word_count", 6:7) %>%
  group_by(year) %>%
  mutate(songs_per_year = n()) %>%
  group_by(songs_per_year, topic, kim_kardashian) %>%
  summarise(word_count = sum(word_count, na.rm = TRUE))%>%
  mutate(normalized_word_count = word_count/songs_per_year)%>%
  ggplot(., aes(x=kim_kardashian, y=normalized_word_count, fill=topic)) +
  geom_bar(stat="identity", position="dodge") +
  labs(x = "", y = "Average swearing per song",
       title = " The Kardashian Effect?",
       subtitle= "214 songs in 13 albums since 2004.")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

But this is a weird categorical ordering, no? We can change this!

```{r kim effect reordered, message=FALSE, warning=FALSE}
Kanye_West$kim_kardashian <- factor(Kanye_West$kim_kardashian,                      
                                    levels = c("Before Kim",
                                               "With Kim",
                                               "After Kim"))
Kanye_West %>%
  gather("topic", "word_count", 6:7) %>%
  group_by(year) %>%
  mutate(songs_per_year = n()) %>%
  group_by(songs_per_year, topic, kim_kardashian) %>%
  summarise(word_count = sum(word_count, na.rm = TRUE))%>%
  mutate(normalized_word_count = word_count/songs_per_year)%>%
  ggplot(., aes(x=kim_kardashian, y=normalized_word_count, fill=topic)) +
  geom_bar(stat="identity", position="dodge") +
  labs(x = "", y = "Average swearing per song",
       title = " The Kardashian Effect?",
       subtitle= "214 songs in 13 albums since 2004.")+
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3,
                                        linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
```

## Lamar's Repertoire

```{r lamar cleaning, include=FALSE, message=FALSE, warning=FALSE}
# The function "function()" creates a function; in this case, our function cleans the text in multiple ways.
cleanCorpus<-function(corpus, customStopwords){
  corpus <- tm_map(corpus, content_transformer(qdapRegex::rm_url))
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, removeWords, customStopwords)
  return(corpus)
}

stops <- c(stopwords('SMART'), "dont", "youre")
# stopwords ('SMART') is a vector with common words (a, the, about...) that we want to remove from the songs

Kendrick_Lamar <- american_rappers %>%
  filter(rapper=="Kendrick Lamar") # creating a dataset only for Kendrick
  
# the next few lines perform a few operations to create a dataset counting appearances of words in all songs.
lamarCorpus_t <- VCorpus(VectorSource(Kendrick_Lamar$lyrics))
lamarCorpus_t <- cleanCorpus(lamarCorpus_t, stops)
LamarTDM_t  <- TermDocumentMatrix(lamarCorpus_t)
lamar_dtm_t  <- DocumentTermMatrix(lamarCorpus_t)
# create a document term matrix

LamarTDMm_t <- as.matrix(LamarTDM_t)
LamarSums_t <- rowSums(LamarTDMm_t)
LamarFreq_t <- data.frame(word=names(LamarSums_t),frequency=LamarSums_t)
rownames(LamarFreq_t) <- NULL
topWords_t  <- subset(LamarFreq_t, LamarFreq_t$frequency >= 50)
#here we are subsetting to those that appear more than 50 times.

topWords_t  <- topWords_t[order(topWords_t$frequency, decreasing=F),]
topWords_t$word <- factor(topWords_t$word, levels=unique(as.character(topWords_t$word)))
# getting the top words in frequency
```

```{r lamar plot, message=FALSE, warning=FALSE}
ggplot(topWords_t, aes(x=word, y=log(frequency))) +
  geom_bar(stat="identity", fill='gold') +
  coord_flip()+
  geom_text(aes(label=frequency), colour="black",hjust=1.25, size=3.0)+
  labs( x="", y= "", title= "Most frequent words in Lamar's repertoire",
        subtitle="All albums since 2011")+
  theme(panel.background = element_rect ("white", "black", .5, "solid"),
        panel.grid.major = element_line(color="grey", size=0.3, linetype= "solid"),
        axis.text = element_text(color="black", size=10),
        title = element_text(color="black", size=10, face="bold"),
        axis.text.x=element_blank(),
        plot.subtitle = element_text(color="black", size=9, face= "plain"),
        legend.position = "bottom")
```
